Management induced changes of soil organic carbon on global croplands

### 2020-12-11 SOCbudgetPaper/SOCbudget.Rmd
update text and style files

---
title: Management induced changes of soil organic carbon on global croplands 
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Kristine
    surname: Karstens
    affiliation: 1
    email: kristine.karstens@pik-potsdam.de
    corresponding: true
  - given_name: Benjamin Leon
    surname: Bodirsky
    affiliation: 1
  - given_name: Your
    surname: Name
    affiliation: 2
  - given_name: Alexander
    surname: Popp
    affiliation: 1
affiliation:
  - code: 1
    address: Potsdam-Institut of Climate Impacts Research, Potsdam, Germany
  - code: 2
    address: Your affiliation 
abstract: |
  Soil organic carbon (SOC) is one of larges carbon stocks on Earth. It is three times larger than the vegetation pool and more than twice as big as the atmospheric pool, when looking into the first meter of the earth soil profile. Human cropping activities led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions. While SOC models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated on the global scale. Here we create a spatial explicit data set for agricultural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils. We estimate that due to arable farming soils have lost over 37 GtOC of which 4 GtOC have been regained within the period 1975--2010. We show that, our results on global scale based on Tier 2 IPCC methodology are in good agreement with Tier 1 default assumptions. We also find that SOC is very sensitive to management decision such as residue recycling indicating the necessity to incorporated better management data in soil models.
bibliography: SOCbudget2.bib
running:
  title: R Markdown Template for Copernicus
  author: Nüst et al.
competinginterests: |
  The authors declare no competing interests.
# OPTIONAL:
algorithms: true
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed.
copyrightstatement: |
  The author's copyright for this publication is transferred to institution/company.
availability:
  codedata: |
    Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/. Data used for the output can be found under <something-with-zenodo>.  
authorcontribution: |
  Karstens wrote code and paper build on work of Bodirsky (and <here_could_be_your_names>). 
  Bodirsky, <here_could_be_your_names> and Popp revised paper.
disclaimer: |
  We like Copernicus.
acknowledgements: |
  Thanks to the rticles contributors!
output:
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
    md_extensions: +raw_attribute
  md_document:
    variant: markdown_github
  bookdown::word_document2: default
  rticles::copernicus_article: default
  bookdown::pdf_document2: default
---
\newpage

@@

```{r child = 'appendix.Rmd'}
```
\newpage

>>>>

---
title: Management induced changes of soil organic carbon on global croplands 
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Kristine
    surname: Karstens
    affiliation: 1
    email: kristine.karstens@pik-potsdam.de
    corresponding: true
  - given_name: Benjamin Leon
    surname: Bodirsky
    affiliation: 1
  - given_name: Your
    surname: Name
    affiliation: 2
  - given_name: Alexander
    surname: Popp
    affiliation: 1
affiliation:
  - code: 1
    address: Potsdam-Institut of Climate Impacts Research, Potsdam, Germany
  - code: 2
    address: Your affiliation 
abstract: |
  Soil organic carbon (SOC) is one of larges carbon stocks on Earth. It is three times larger than the vegetation pool and more than twice as big as the atmospheric pool, when looking into the first meter of the earth soil profile. Human cropping activities led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions. While SOC models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated on the global scale. Here we create a spatial explicit data set for agricultural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils. We estimate that due to arable farming soils have lost over 37 GtOC of which 4 GtOC have been regained within the period 1975--2010. We show that, our results on global scale based on Tier 2 IPCC methodology are in good agreement with Tier 1 default assumptions. We also find that SOC is very sensitive to management decision such as residue recycling indicating the necessity to incorporated better management data in soil models.
bibliography: SOCbudget2.bib
running:
  title: R Markdown Template for Copernicus
  author: Nüst et al.
competinginterests: |
  The authors declare no competing interests.
# OPTIONAL:
algorithms: true
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed.
copyrightstatement: |
  The author's copyright for this publication is transferred to institution/company.
availability:
  codedata: |
    Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/. Data used for the output can be found under <something-with-zenodo>.  
authorcontribution: |
  Karstens wrote code and paper build on work of Bodirsky (and <here_could_be_your_names>). 
  Bodirsky, <here_could_be_your_names> and Popp revised paper.
disclaimer: |
  We like Copernicus.
acknowledgements: |
  Thanks to the rticles contributors!
appendix:
output:
  rticles::copernicus_article: default
  md_document:
    variant: markdown_github
  bookdown::word_document2: default
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
    md_extensions: +raw_attribute
  bookdown::pdf_document2: default
---
\newpage

@@

```{r child = 'appendix.Rmd'}
```

### 2020-12-11 SOCbudgetPaper/appendix.Rmd
update text and style files

```{r regSOCtable, out.width = "100%", fig.cap = captionregSOCtable, echo=FALSE, fig.pos="H"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/reg_comparisonfigure.png")
```

@@

```{r initSOC, out.width = "100%", fig.cap = captioninitSOC, echo=FALSE, fig.pos="H"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_lu.png")
```

>>>>

```{r regSOCtable, out.width = "100%", fig.cap = captionregSOCtable, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/reg_comparisonfigure.png")
```

@@

```{r initSOC, out.width = "100%", fig.cap = captioninitSOC, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_lu.png")
```

### 2020-12-11 SOCbudgetPaper/discussion.Rmd
update text and style files

According to [@sanderman_soil_2017] cumulative SOC emissions since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing. Our estimate of 37 GtC in 2010 for cropland emissions only is twice as high as [@sanderman_soil_2017] estimations. However, there are large uncertainties in modelling SOC at the global scale, and the [@sanderman_soil_2017] results have been pointed out as potentially conservatively low compared with experimental results, leaving our results in an aceptable range.

Furthermore, [@sanderman_soil_2017] modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all. [@pugh_simulated_2015] considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account. The study moreover concludes that crop productivity gains (increasing yield levels by 18%) do not lead to a substantial decrease in SOC emissions, accounting for only less than 1% change in SOC emissions.

Our study for the first time uses a dynamic management dataset as driver for SOC dynamics. We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been out weighted by improved agricultural yields and practices. This challenges [@pugh_simulated_2015] findings of only small effects due to improved practices.

Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates play the most important role, followed by (2) improved manure recycling (e.g. due to improved animal waste management systems) and (3) the adoption of no tillage practices.

Modelling management effects on the global scale comes however with parametric and structural uncertainties. As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities. [@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains. Moreover, the study pointed out that plant breeding shifts allometries which might not be reflected in outdated data sources. Following this line of argumentation, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding. However, with an evaluation of management effects (see below), there is no indication for a systematic overrating of residue biomass. Rather, it is more likely that we are still missing carbon inputs to the soil.

@@

To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (@ipcc_2006_2006, @ipcc_2019_2019) which are based on measurement data for croplands (see \@ref(tab:SCFtable)). To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones.

```{r captionSCFtable, echo=FALSE}
captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation. We compare average values for four IPCC climate zone classifications. IPCC 2006 and 2019 default factors (medium input) as well as altered factors for low input of organic matter without any other subsystem consideration are compared to results for 1990 and 2010 of this study (SOC budget)."
```

@@

Our estimates correspond very well to the default stock-change factors used for the tier 1 estimation of IPCC, 2006. For the tropical regions the assumptions changed notably from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019. Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems might be an overestimation of actual SOC state. The additional effect of considering low inputs of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for 14-21% of it.

With regard to the time trend, our study shows the substantial impact of changing management factors on the development of $F_{SCF}$.

@@

The world's SOC stock and its changes are highly uncertain (cite), which is seen in the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).

@@

```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="H"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/glo_comparisonfigure.png")
```

The global estimates of SOC stock from this study are on the lower end compared to other modelled results or more data driven estimates. Looking at regional results \@ref(append:regcompare), our estimates turn out to be in good agreement for most regions, with the largest deviations for boreal areas. Considering that the model was parametrized for croplands, these mismatches are not suprising as the temperature effects on decomposition are fundamentally different for permafrost soils. To avoid that this bias influences our results, our study focuses exclusively on cropland soils, excluding most of the boreal zone. Moreover, when focussing on SOC changes on cropland, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.

Our estimates for total SOC stocks of the world, as well as our SOC initialization, are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner. For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes. This leaves carbon inputs and decay behavior for natural land and pastures rather uncertain. The absolute values of stocks and emissions from land-use change therefore have to be used with caution. Especially in less forested areas the natural land representation higher uncertainty may arise from the parametrization assumptions of natural litterfall.

Nevertheless, total SOC stocks are in a reasonable range and results for SOC stock changes in relative or absolute numbers are not altered by pristine vegetated areas at all. The SOC gap from changed management is only affected by the natural vegetation representation in the case of land use transitions. In this case, former natural land SOC stocks decline. However, as pointed out before, due to low cropland expansion rates, land use change-related SOC declines are outweighed by the general trend of increasing SOC stocks on existing cropland due to management improvements.
We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parametrization of natural litterfall, which tend to change the global SOC stock substantially for different choices
(see appendix). This shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.

## Important shortcomings --> might be added as well

The **initialization of SOC stocks** in 1960 is assumed to be in steady state considering the land use pattern and management effects of the initialization year followed by a run-in period of 15 years, till the analysis of the results start. That is inline with IPCC guidelines calling for a run-in period of 5-20 years. However, the results are heavily influenced by it as \@ref(append:initcompare) suggested. Not only the level of the SOC gap is shifted, also the declining trend is much smaller or even not visible at all anymore. For the artificial assumption of starting with potential natural vegetation in 1961 the SOC gap is fluctuating around 10 GtC for the period 1975--2010. For the more moderate assumption, that half of the transition towards the steady-state under land use has already appeared in 1960, the SOC gap is around 23 GtC in 2010. The relative declining trend for the SOC gap between 1975 and 2010 (-12%) is however almost the same as for our historic estimation (-14%). According to LUH [@hurtt_harmonization_2020] cropland area has increased over 62% since 1900 leaving our choice of initialization values too low, since reaching a new steady-state can take decades or even centuries. Having no yield or management data available before our initialization year leaves us with imperfect initialization knowledge, where our initialization choice is probably closer to real world values than starting from a natural vegetated steady-state.

Following the IPCC guidelines this study has limited its focus to the **first 30 cm of the soil profile**. In this regard several aspects are over simplified within our approach. Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil. This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas. Consequently, our SOC loss estimates are too high.
Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems. It has been argued that for intensively tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers. Following this arguments SOC stocks in croplands might even be underestimated.

Carbon displacement via **leaching and erosion** is neglected in this study. As pointed out by several studies (cite), the final fate of leached or eroded carbon is uncertain. Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on SOC emissions, displaced but not emitted SOC can be treated as SOC stayed on croplands.

>>>>

According to [@sanderman_soil_2017] cumulative SOC emissions since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing. It was also pointed out that these results might be conservatively low compared to experimental results. Considering the huge uncertainties in modelling SOC on the global scale, our estimate of 37 GtC in 2010 for cropland emissions only still seems consistent with [@sanderman_soil_2017] estimates.

Furthermore, the results of [@sanderman_soil_2017] modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all. [@pugh_simulated_2015] considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account. The study moreover concludes that productivity gains for crops while increasing yield levels by 18%, do not spill over to a substantial decrease in SOC emissions, but only account for less than 1% change in SOC emissions.

Our study for the first time uses a dynamic management dataset as driver for SOC dynamics. We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been out weighted by improved agricultural yields and practices. This is challenging [@pugh_simulated_2015] findings of only small effects due to improved practices.
Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates, (2) improved manure recycling (e.g. due to improved animal waste management system) and (3) the adoption of no tillage practices.

Modelling management effects on the global scale comes however with parametric and structural uncertainties. As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities. [@keel_large_2017] questioned whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains. Moreover, the study pointed out that plant breeding shift allometries and might be not reflected in outdated data sources. Following this argumentation SOC results shown in this study might overestimate actual SOC stocks due to an overestimation in residue biomass, which ratio compared to the harvested organ is normally reduced due to breeding. However, looking into management effect evaluation (see below) there is no indication for a systematic overrating of residue biomass, more likely we are still missing carbon inputs to the soil.

@@

To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (@ipcc_2006_2006, @ipcc_2019_2019) which are based on measurement data for croplands (see \@ref(tab:SCFtable)). To allow for comparison, we aggregated our stock change factors to the four IPCC climate zones.

```{r captionSCFtable, echo=FALSE}
captionSCFtable <- "This table shows different estimates of stock change factors $F^{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation. We compare average values for four IPCC climate zone classifications. IPCC 2006 and 2019 default factors (medium input) as well as altered factors for low input of organic matter without any other subsystem consideration are compared to results for 1990 and 2010 of this study (SOC budget)."
```

@@

Our estimates correspond very well to the default stock-change factors used for the tier 1 estimation of IPCC, 2006. For the tropical regions the assumptions changed notably from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019. Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems, might be an overestimation of actual SOC state. The additional effect of considering low input of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for at least 5-7% of it.

Looking again on the time trend our study shows the substantial impact of changed management on the development of $F_{SCF}$.

@@

The world's SOC stock and its changes are highly uncertain (cite), which is visible by the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).

@@

```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/glo_comparisonfigure.png")
```

The global estimates of SOC stock by this study are on the lower end compared to other modelled results or more data driven estimates. Looking on regional results \@ref(append:regcompare), our estimates turn out to be in good agreement for most regions with the largest deviations for boreal areas. Considering that the model was parametrized for croplands, these mismatches are not surprising since the temperature effects on decomposition are fundamentally different for permafrost soils. To avoid that this bias influences our results, our study focuses exclusively on cropland soils, excluding most of the boreal zone. Moreover, when focussing on SOC changes, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.

Our estimates for total SOC stocks of the world, as well as our SOC initialization are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner. For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes. This leaves carbon inputs and decay behavior for natural land and pastures rather uncertain. The absolute values of stocks and emissions from land-use change therefore have to be used with caution. Especially in less forested areas the natural land representation might be off, due to parametrization assumptions of the natural litterfall.

Nevertheless, total SOC stocks are in a reasonable range and results on SOC stock changes in relative or absolute numbers are not altered by pristine vegetated areas at all. Moreover, the SOC gap from changed management is only affected from the natural vegetation representation in case of land use transitions. Then former natural land SOC stocks are declining to lower levels. However, as pointed out before, due to low cropland expansion rates, their SOC declines are outweighed by the general trend of increasing SOC stocks on existing cropland due to management improvements.
We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parametrization of natural litterfall, which tend to change the global SOC stock substantially for different choices
(see appendix). It shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.

## Important short comings --> might be added as well

The initialization of SOC stock in 1960 is assumed to be in steady state followed by a run-in period of 15 years, till the analysis of the results start. That is inline with IPCC guidelines calling for a run-in period of 5-20 years. However, the results in the early years might be still distorted by a longer living memory effects than this 15 years of spin-up.

Following the IPCC guidelines this study has limited its focus to the first 30 cm of the soil profile. In this regard several aspects are over simplified within our approach. Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil. This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas. Consequently, our SOC loss estimates are too high.
Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems. It has been argued that for intensively tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers. Following this argumentation SOC stocks in croplands might even be underestimated.

Carbon displacement via leaching and erosion is neglected in this study. As pointed out by several studies (cite), the final fate of leached or eroded carbon is uncertian. Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach on SOC focusing especialy on emissions, displaced but not emitted SOC can be treated as SOC stayed on croplands.

### 2020-12-11 SOCbudgetPaper/methods.Rmd
update text and style files

$SOC$ converge towards the calculated steady-state stock $SOC^{\mathrm{eq}$ for each grid cell $i$, each annual time step $t$, each land-use type $lu$ and each sub-pool $sub$ like

@@

Additionally to the steady-state method [@calvo_buendia_ipcc_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]). Here, stocks are calculated via stock change factors ($F^{\mathrm{SCF}}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data. Estimates of $F^{\mathrm{SCF}}$ are differentiated by different crops, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these. Moreover, estimates of $F^{\mathrm{SCF}}$ vary for different climatic zones ($c$) specified by the IPCC (see Fig. \@ref(append:climatemap)). The actual $SOC$ stocks are thus calculated based on a given reference stock $SOC^{\mathrm{ref}$ by

@@

To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for the three drivers: (1) In the \textit{constTillage} scenario the adoption of no-tillage practises are neglected (adoption starts in 1975 according to the available data set); (2) the \textit{constResidues} and the \textif{constManure} scenario assume constant input rates from residues resp. manure (in $\tfrac{t}{ha}$) at the level of 1975 onwards. Within the \textit{constResidue} scenario at least three effects overlay each other: (1) yields and with them residue biomass increase due to productivity gains, (2) rates of residue left or returned to fields are raising, and (3) shifts of crop groups change the amount of residue biomass due to crop-group specific harvest index values. The \textit{constManagement} combines all three scenarios.

@@

 \begin{table*}[h]
 \caption{Scenarios for sensitivity analysis on litter parameterization}
 \begin{tabular}{l l l}
 \tophline
  Scenario name & plant type & n/c & lg/c \\
  \middlehline
  \textit{LitterPNV-CenturyAverage} & mix over all tree types & 0.015  & 0.25 \\
  \textit{LitterPNV-PerennialGrasses} & perennial grasses     & 0.029  & 0.11 \\
  \textit{LitterPNV-AASS}  & Arctic alpine subshrub           & 0.022  & 0.22 \\
  \textit{LitterPNV-BLEAF} & Temperate broadleaf evergreen    & 0.013  & 0.27 \\
  \textit{LitterPNV-BRCF}  & Boreal coniferous forest         & 0.013  & 0.17 \\
  \textit{LitterPNV-MTCF}  & Maritime coniferous forest       & 0.0094 & 0.22 \\
  \textit{LitterPNV-TREF}  & Tropical evergreen forest        & 0.016  & 0.30 \\
 \bottomhline
 \end{tabular}
 \belowtable{}
 (\#tab:scen_litterpnv)
 \end{table*}
```

>>>>

$SOC$ converge towards the calculated steady-state stock $SOC^{\mathrm{eq}}$ for each grid cell $i$, each annual time step $t$, each land-use type $lu$ and each sub-pool $sub$ like

@@

Additionally to the steady-state method [@calvo_buendia_ipcc_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]). Here, stocks are calculated via stock change factors ($F^{\mathrm{SCF}}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data. Estimates of $F^{\mathrm{SCF}}$ are differentiated by different crops, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these. Moreover, estimates of $F^{\mathrm{SCF}}$ vary for different climatic zones ($c$) specified by the IPCC (see Fig. \@ref(append:climatemap)). The actual $SOC$ stocks are thus calculated based on a given reference stock $SOC^{\mathrm{ref}}$ by

@@

To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for the three drivers: (1) In the \textit{constTillage} scenario the adoption of no-tillage practises are neglected (adoption starts in 1975 according to the available data set); (2) the \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp. manure (in $\tfrac{t}{ha}$) at the level of 1975 onwards. Within the \textit{constResidue} scenario at least three effects overlay each other: (1) yields and with them residue biomass increase due to productivity gains, (2) rates of residue left or returned to fields are raising, and (3) shifts of crop groups change the amount of residue biomass due to crop-group specific harvest index values. The \textit{constManagement} combines all three scenarios.

@@

 \begin{table*}[h]
 \caption{Scenarios for sensitivity analysis on litter parameterization}
 \begin{tabular}{l l l l}
 \tophline
  Scenario name & plant type & n/c & lg/c \\
  \middlehline
  \textit{LitterPNV-CenturyAverage} & mix over all tree types & 0.015  & 0.25 \\
  \textit{LitterPNV-PerennialGrasses} & perennial grasses     & 0.029  & 0.11 \\
  \textit{LitterPNV-AASS}  & Arctic alpine subshrub           & 0.022  & 0.22 \\
  \textit{LitterPNV-BLEAF} & Temperate broadleaf evergreen    & 0.013  & 0.27 \\
  \textit{LitterPNV-BRCF}  & Boreal coniferous forest         & 0.013  & 0.17 \\
  \textit{LitterPNV-MTCF}  & Maritime coniferous forest       & 0.0094 & 0.22 \\
  \textit{LitterPNV-TREF}  & Tropical evergreen forest        & 0.016  & 0.30 \\
 \bottomhline
 \end{tabular}
 \belowtable{}
 (\#tab:scenlitterpnv)
 \end{table*}
```

### 2020-12-11 SOCbudgetPaper/results.Rmd
update text and style files

```{r SOCmaps, out.width = "18cm", fig.cap = caption_SOCmaps, echo=FALSE, fig.pos="H"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/4panelfigure.png")
```

@@

```{r SOCscen, out.width = "18cm", fig.cap = caption_SOCscen, echo=FALSE, fig.pos="H"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/scenario.png")
```

@@

```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE, fig.pos="H"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/OuFlowFig.png")
```

>>>>

```{r SOCmaps, out.width = "18cm", fig.cap = caption_SOCmaps, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/4panelfigure.png")
```

@@

```{r SOCscen, out.width = "18cm", fig.cap = caption_SOCscen, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/scenario.png")
```

@@

```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/OuFlowFig.png")
```
